Annotations on Quantum Pseudorandomness and $k$-Designs

### 2020-11-22 PHYS790/final.tex
Added a bit on the comments paper.

Normalizing this sum allows us to view it as a \emph{probability distribution}. Furthermore, the evolution of our random circuit scheme is amenable to Markov chain analysis. As an illustrative example, suppose we have an $n$-qubit system and so far have applied $t$ steps of our scheme. As per usual, we pick our $2$-qubit unitary $U \sim \mathbb{U}(4)$. Now if we were to pick qubits $1,2$, our random $2$-qubit unitary would only act on the first two qubits non-trivially whereas every other qubit would just transform via the identity map. By the explicit calculation shown in Equation \ref{concreteT}, the expected \emph{coefficients} for step $t+1$ will work out to be

@@

For this case, we will just sketch the overarching ideas of the section and leave some details to the appendix of this report and \cite{harrlow}. Once again, we restrict our attention to the coefficients of diagonal states and delete the identity $\sigma_0 \otimes \sigma_0$ in order for the chain to become irreducible. The analysis starts with the consideration of an auxiliary chain called the \emph{zero chain}. The zero chain describes the number of zeros present in a string. The idea is that the number of zeroes slow the mixing process since picking two zeroes effectively changes nothing. If we can show that distribution of zeros mixes well-enough, it can be used to further infer about the mixing of the total chain.
The transition probabilities are worked out in a lemma
%
\begin{lemma} \label{zerochain}
  The zero chain has the following transition matrix $P$ on the state space $\Omega = \{1,2, \cdots,n\}$. This chain actually counts the number of non-zeros of state.
  \begin{equation}
    P(x,y) = \begin{cases}
    1 - \frac{2x(3n-2x -1)}{5n(n-1)} & y = x \\
    \frac{2x(x-1)}{5n(n-1)} & y = x - 1 \\
    \frac{6x(n-x)}{5n(n-1)} & y = x + 1 \\
    0 & \text{otherwise}
    \end{cases}
  \end{equation}
  for $1 \leq x,y \leq n$
\end{lemma}
\begin{proof}
  This can be derived through a simple counting argument. For example, for the number of zeros to decrease ($x$ increases), two sites with a non-zero value and a zero value must be chosen and the chain must transition to pair of non-zero values. $9$ out of the $15$ possible transition values consist of two non-zero values. There the probabilty of picking a zero and a non-zero value is $\frac{2x(n-x)}{n(n-1)}$, bringing the total value to $\frac{9}{15}\frac{2x(n-x)}{n(n-1)} = \frac{6x(n-x)}{5n(n-1)}$.
\end{proof}
% Mention stationary distribution
%
\noindent It is also simple to show that the stationary distribution $\pi_0$ of the zero chain will be
%
\begin{equation}
    \pi_0(x) = \frac{3^x{n \choose x}}{4^n - 1}
\end{equation}
%
The integral property about this zero chain is that it mixes rapidly
%
\begin{theorem} \label{zerochainmixing}
  The zero chain mixes in $\Theta(n\log\frac{n}{\epsilon})$ time.
\end{theorem}
%
\noindent which leads to the rapid mixing of the total chain
\begin{theorem} \label{totalchainmixing}
  The total chain mixes in time $\Theta(n \log\frac{n}{\epsilon})$
\end{theorem}
\begin{remark}
  Diniz and Johnathan point out in \cite{diniz} a faulty assumption made by the authors of \cite{harrlow}. In the proof of Corollary \ref{totalchainmixing},
\end{remark}
The proofs of these two theorems will also be assigned a portion of the appendix.
By the discussion of Corollary \ref{muchainmix}, this proves Lemma \ref{mainlemma}.

>>>>

Normalizing this sum allows us to view it as a \emph{probability distribution}. Furthermore, the evolution of our random circuit scheme is amenable to Markov chain analysis. As an illustrative example, suppose we have an $n$-qubit system and so far have applied $t$ steps of our scheme. As per usual, we pick our $2$-qubit unitary $U \sim \mathbb{U}(4)$. Now if we were to pick qubits $1,2$, our random $2$-qubit unitary would only act on the first two qubits non-trivially whereas every other qubit would just transform via the identity map. By the explicit calculation shown in Equation \ref{concreteT} and linearity, the expected \emph{coefficients} for step $t+1$ will work out to be

@@

For this case, we will just sketch the overarching ideas of the section and leave some details to the appendix of this report and \cite{harrlow}. Once again, we restrict our attention to the coefficients of diagonal states and delete the identity $\sigma_0 \otimes \sigma_0$ in order for the chain to become irreducible. The analysis starts with the consideration of an auxiliary chain called the \emph{zero chain}. The zero chain describes the number of zeros present in a string. The idea is that the number of zeroes slow the mixing process since picking two zeroes effectively changes nothing. If we can show that distribution of zeros mixes well-enough, it can be used to further infer about the mixing of the total chain.
The transition probabilities are worked out in a lemma
%
\begin{lemma} \label{zerochain}
  The zero chain has the following transition matrix $P$ on the state space $\Omega = \{1,2, \cdots,n\}$. This chain actually counts the number of non-zeros of state.
  \begin{equation}
    P(x,y) = \begin{cases}
    1 - \frac{2x(3n-2x -1)}{5n(n-1)} & y = x \\
    \frac{2x(x-1)}{5n(n-1)} & y = x - 1 \\
    \frac{6x(n-x)}{5n(n-1)} & y = x + 1 \\
    0 & \text{otherwise}
    \end{cases}
  \end{equation}
  for $1 \leq x,y \leq n$
\end{lemma}
\begin{proof}
  This can be derived through a simple counting argument. For example, for the number of zeros to decrease ($x$ increases), two sites with a non-zero value and a zero value must be chosen and the chain must transition to pair of non-zero values. $9$ out of the $15$ possible transition values consist of two non-zero values. There the probabilty of picking a zero and a non-zero value is $\frac{2x(n-x)}{n(n-1)}$, bringing the total value to $\frac{9}{15}\frac{2x(n-x)}{n(n-1)} = \frac{6x(n-x)}{5n(n-1)}$.
\end{proof}
% Mention stationary distribution
%
\noindent It is also simple to show that the stationary distribution $\pi_0$ of the zero chain will be
%
\begin{equation}
    \pi_0(x) = \frac{3^x{n \choose x}}{4^n - 1}
\end{equation}
%
The integral property about this zero chain is that it mixes rapidly
%
\begin{theorem} \label{zerochainmixing}
  The zero chain mixes in $\Theta(n\log\frac{n}{\epsilon})$ time.
\end{theorem}
%
\noindent which leads to the rapid mixing of the total chain
\begin{theorem} \label{totalchainmixing}
  The total chain mixes in time $\Theta(n \log\frac{n}{\epsilon})$
\end{theorem}
\begin{remark}
  Diniz and Johnathan point out in \cite{diniz} a faulty assumption made by the authors of \cite{harrlow}. In the proof of Corollary \ref{totalchainmixing}, Harrow and Low make the argument that once every site has been paired with another site with non-zero value, the total chain will be uniformly mixed. This would reduce the mixing time to a coupon collectors problem. However, the authors of \cite{diniz} show a simple counterexample to this statement. Let us focus our attention to a $3$-qubit system with initial state $(0,0,1)$. We want to see the resulting distribution after hitting all three sites with two transitions.
  First, observe that we have the following transitions $(0,0,1) \rightsquigarrow (1,0,0)$
  \begin{align*}
    & 001 \rightsquigarrow 100 \rightsquigarrow 100 \\
    & 001 \rightsquigarrow 010 \rightsquigarrow 100 \\
    & 001 \rightsquigarrow 001 \rightsquigarrow 100
  \end{align*}
  whereas we have the transitions $(0,0,1) \rightsquigarrow (0,0,1)$
  \begin{align*}
    & 001 \rightsquigarrow 001 \rightsquigarrow 001 \\
    & 001 \rightsquigarrow 001 \rightsquigarrow 001 \\
  \end{align*}
   by first hitting the sites $2,3$ and $1,3$. The second transition hits the sites in the reverse order.
\end{remark}

The proofs of these two theorems will also be assigned a portion of the appendix.
By the discussion of Corollary \ref{muchainmix}, this proves Lemma \ref{mainlemma}.
